---
title: "Untitled"
author: "Sateesh"
date: "3/23/2020"
output: pdf_document
---
```{r}
if(!require("pacman")) install.packages("pacman")
pacman::p_load(tidyverse,readxl,shiny,ggplot2,lubridate,shinydashboard,shiny,DT,plotly,rworldmap)
```
#Reading the Data
```{r}
ebola <- read.csv('ebola_2014_2016.csv')
ebola$Date<- as.Date(as.character(ebola$Date),format = "%m/%d/%Y")
```
#Data wrangling

##Removing the duplicate records of same day and from same country
```{r}
ebola1 <- ebola %>% group_by(ebola$Country,ebola$Date) %>% summarise(max(No..of.suspected.cases),max(No..of.probable.cases),max(No..of.confirmed.cases),max(No..of.confirmed..probable.and.suspected.cases),max(No..of.suspected.deaths),max(No..of.probable.deaths),max(No..of.confirmed.deaths),max(No..of.confirmed..probable.and.suspected.deaths))
```
```{r}
ebola1 <- data.frame(ebola1)
names(ebola1) <- names(ebola)
```
```{r}
ebola <- ebola1
```

```{r}
ebola[order(ebola$Country,ebola$Date),]
```
#Parameters to be passed to the input controls
```{r}
parameters <- colnames(ebola)
parameters <- parameters[c(3:10)]
```

1.Daily increase of cases for each country:

```{r}
ggplot(data = ebola[!is.na(ebola$No..of.confirmed.deaths)& ebola$Country == "Guinea",],aes(Date,No..of.confirmed.deaths))+
  geom_point()+
  geom_line()
```
2. Finding average number of cases each month for every country.
```{r}
df <- ebola
df$month <- months(ebola$Date)
df$year <- year(ebola$Date)
df$month <- factor(df$month ,levels = c("January","February","March","April","May","June","July","August","September","October","November","December"))
df$year <- factor(df$year)
```

```{r}
str(df)
```

```{r}
df <- df %>% group_by(df$Country,df$month,df$year) %>% summarise(mean(No..of.suspected.cases),mean(No..of.probable.cases),mean(No..of.confirmed.cases),mean(No..of.confirmed..probable.and.suspected.cases),mean(No..of.suspected.deaths),mean(No..of.probable.deaths),mean(No..of.confirmed.deaths),mean(No..of.confirmed..probable.and.suspected.deaths))
df <- data.frame(df)
```
```{r}
names <- colnames(ebola)
colnames(df) <- c("Country","Month","Year",names[3:10])
df[,4:11]<- round(df[,4:11],0)
```

# Monthly increase graph
```{r}
ggplot(data = df[!is.na(df$No..of.confirmed.deaths) & df$Country == "Guinea" ,] , aes(x = Month , y = No..of.confirmed.deaths , group = 1))+
  geom_point()+
  geom_line()+
  facet_wrap(~Year)+
  theme(axis.text.x=element_text(size=8, angle=90))
```
## Position of each country at the end of the month
```{r}
df1 <- df %>% group_by(df$Country,df$Month,df$Year) %>% summarise(max(No..of.confirmed.cases),max(No..of.confirmed..probable.and.suspected.cases),max(No..of.confirmed.deaths),max(No..of.confirmed..probable.and.suspected.deaths))
df1 <- data.frame(df1)
str(df1)
```
Cumulative number of cases
```{r}
df1 <- df1[,c(1,2,3,5,7)]
names(df1) <- c("Country","Month","Year","Cumulative cases","Cumulative deaths")
df1[,c(4,5)] <- round(df1[,c(4,5)])
```
## Adding Country Codes to plot world map
```{r}
df2 <- data.frame("Country" = c("Liberia","United States of America","Guinea","Italy","Mali","Nigeria","Senegal","Sierra Leone","Spain","United Kingdom"),"ISO3V10" = c("LBR","USA","GIN","ITA","MLI","NGA","SEN","SLE","ESP","GBR"))

df3 <- df1 %>% left_join(df2,by = "Country")
```

Shiny Dashboard # Daily cases graph & Number of cases and death after each month

```{r}
ui <- dashboardPage(
  dashboardHeader(title = "Ebola cases"),
  dashboardSidebar(sidebarMenu(
      menuItem("Dashboard", tabName = "dashboard", icon = icon("dashboard")),
      menuItem("Cases on World Map", tabName = "widgets", icon = icon("th"))
    )),
  dashboardBody(
    tabItems(
      # First tab content
      tabItem(tabName = "dashboard",
    fluidRow(
      box(
        plotlyOutput("plot1",height = 300)
      ),
      box(
        title = "Controls",
         div(style = 'overflow-y: scroll', 
        selectInput(inputId = "country",
                             label = "Country",
                             choices = ebola$Country,
                             selected = ""
                             
                 ),
        selectInput(inputId = "parameter",
                             label = "Parameter",
                             choices = parameters,
                             selected = ""
                             
                 ),
        selectInput(inputId = "month",
                             label = "Month",
                             choices = df1$Month,
                             selected = ""
                             
                 ),
        selectInput(inputId = "year",
                             label = "Year",
                             choices = df1$Year,
                             selected = ""
                             
                 )
        
        
      )),
       box(
        #div(style = 'overflow-x: scroll', tableOutput('table1')),
        dataTableOutput("table1")
      ),
      box(
        plotlyOutput("plot2")
      )
      
    )
  ),
  tabItem(tabName = "widgets",
          h2("World wide cases"),
    fluidRow(
      box(
        plotOutput("plot5")),
        box(
        title = "Controls",
         div(style = 'overflow-y: scroll', 
        selectInput(inputId = "month1",
                             label = "Month",
                             choices = df3$Month,
                             selected = ""
                             
                 ),
        selectInput(inputId = "year1",
                             label = "Year",
                             choices = df3$Year,
                             selected = ""
                             
                 ),
        selectInput(inputId = "parameter1",
                             label = "Parameter",
                             choices = c("Cumulative cases","Cumulative deaths"),
                             selected = ""
                             
                 )
        
        
      )
      )
    )
  )

  
  
  )
   

)
)

server <- function(input, output) {
 output$plot1 <- renderPlotly(
    ggplot(data = ebola[ !is.na(input$parameter)&ebola$Country == input$country,],aes(x = Date))+
      aes_string(y = input$parameter)+
      geom_point()+
      geom_line()+
      ggtitle("Daily Increase in cases of selected country in selected year")+
      ylab("No of Cases")
  )
  output$table1 <- DT::renderDataTable(
  df1[df1$Month == input$month & df1$Year == input$year ,] %>% arrange(desc(`Cumulative cases`)) %>% select(-Month,-Year),
  caption = "Position of each country at Selected Month and Year"
  )
  output$plot2 <- renderPlotly(
    ggplot(data = df[!is.na(input$parameter)&df$Country == input$country & df$Year == input$year,],aes(x = Month))+
      aes_string(y = input$parameter)+
      geom_bar(stat = "identity", fill = "cyan")+
      ggtitle("Monthly Average cases of selected Country in selected Year")+
      ylab("No of cases")+
      theme(plot.title = element_text(size = 10, face = "bold"))
  )
 sPDF <- reactive(joinCountryData2Map(df3[df3$Year == input$year1 & df3$Month == input$month1,]
              , joinCode = "ISO3"
              , nameJoinColumn = "ISO3V10"
     ))
  output$plot5 <- renderPlot(
    mapParams <- mapPolys(sPDF()
              , nameColumnToPlot=input$parameter1 ,catMethod="fixedWidth" , missingCountryCol	
= 'darkgrey' , mapTitle = input$parameter1
              )
  )
}

shinyApp(ui, server)
```

